---
title: "iat"
output: html_document
date: "2025-10-04"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IAT

This is our amazing IAT cleaning document.

## Setup

First of all we use the tidyverse package to read our CSV into a data table.

```{r csv, message = FALSE}
library(tidyverse)
dataset <- read_csv("data/iat_raw.csv", show_col_types = FALSE)
```

Then we can filter all columns except those interesting to us using `subset()`. Using `c()` we can build (or "combine") a list of columns that we want to keep.

```{r filter}
dataset <- subset(dataset, select = c(
  "private.id",
  "condition",
  "trial.num",
  "reaction.time",
  "correct",
  "block"
))
```

## The Algorithm

We pre-processed the data in Excel to fulfill steps 1-3.

### Step 4

Remove any trials with more than 10,000ms latency. We can use the subset function again here, keeping all rows where the "Reaction Time" field is <= 10,000ms.

```{r step4}
step4 <- subset(dataset, `reaction.time` <= 10000)
```

```{r step4view, echo = FALSE, results='asis'}
library(knitr)
kable(subset(dataset, `reaction.time` > 10000), caption = "Filtered data")
```

### Step 5

Remove any participants for whom more than 10% of remaining trials have latencies faster than 300ms.

First, we count the number of rows (i.e. trials) belonging to each participant

```{r step5}
trial_counts <- step4 %>% 
  count(`private.id`, name="total")
```

Then, count the number of rows (i.e. trials) where a participant had a latency of < 300ms.

```{r}
too_fast_trials <- step4 %>% 
  group_by(`private.id`) %>% 
  filter(`reaction.time` < 300) %>% 
  count(`private.id`, name="too_fast")
```

Next, we want to find the participants with more than 10% of trials having a latency of < 300ms.

```{r}
under_latency_threshold <- inner_join(trial_counts, too_fast_trials, by = join_by(`private.id`)) %>% 
  mutate(percent_too_fast = too_fast / total * 100) %>% 
  filter(percent_too_fast > 10)
```

To complete this step, we now filter out all participants (and their trials) from our dataset who fall under the 10% latency threshold (calculated above).
To do this we perform an "anti-join", which takes our `step4` table and removes any rows with a `Participant Private ID` contained in our `under_latency_threshold` table (i.e. participants who are too fast, too often)

```{r}
step5 <- anti_join(step4, under_latency_threshold, by = join_by(`private.id`))
```

### Step 6

"correct" column == 0
Blocks A1 & B1

```{r}
training_blocks <- subset(step5, block %in% c("A1", "B1"))
training_blocks

# Group the data by participants AND blocks
training_blocks %>% group_by(`private.id`, block) -> participants_by_training_blocks
participants_by_training_blocks

# Get each participant/block/trial combination that had an incorrect response
participants_by_training_blocks %>% filter(duplicated(`trial.num`)) %>% distinct(`private.id`, `trial.num`, block, .keep_all = TRUE) -> training_errors
training_errors

# Remove all trials that were ever incorrect, *including* their corrected responses
anti_join(participants_by_training_blocks, training_errors, by = join_by(`private.id`, `trial.num`, block)) -> training_only_correct
training_only_correct

test_blocks <- subset(step5, block %in% c("A2", "B2"))
test_blocks

# Group the data by participants AND blocks
test_blocks %>% group_by(`private.id`, block) -> participants_by_test_blocks
participants_by_test_blocks

test_blocks %>% filter(correct == 0) -> test_errors
test_errors

anti_join(participants_by_test_blocks, test_errors, by = join_by(`private.id`, `trial.num`, block)) -> test_only_correct
test_only_correct

all_only_correct <- rbind(training_only_correct, test_only_correct)
all_only_correct

error_penalties <- all_only_correct %>% group_by(`private.id`, block) %>% summarise(error_penalty = mean(`reaction.time`) + (2 * sd(`reaction.time`)))
error_penalties

# Apply the error penalty to training block data
left_join(training_errors, error_penalties, by = join_by(`private.id`, block)) %>% subset(select = -`reaction.time`) %>% rename(`reaction.time` = error_penalty) -> training_errors_recalculated
training_errors_recalculated

# Apply the error penalty to test data
left_join(test_errors, error_penalties, by = join_by(`private.id`, block)) %>% subset(select = -`reaction.time`) %>% rename(`reaction.time` = error_penalty) -> test_errors_recalculated
test_errors_recalculated

all_only_correct <- rbind(all_only_correct, training_errors_recalculated)
all_only_correct <- rbind(all_only_correct, test_errors_recalculated)

step6 <- all_only_correct

```

# Step 7

Calculate the mean reaction time across blocks for each participant.

```{r}
mean_latency_per_block <- step6 %>% group_by(`private.id`, block) %>% summarise(mean_reaction_time = mean(`reaction.time`))
mean_latency_per_block
```

Calculate the mean differences between blocks A1&B1 and A2&B2

```{r}
# Calculate difference between mean latency of B1 and A1
a1_mean_latency <- mean_latency_per_block %>% filter(block %in% "A1") %>% rename(a1 = mean_reaction_time)
b1_mean_latency <- mean_latency_per_block %>% filter(block %in% "B1") %>% rename(b1 = mean_reaction_time)
b1_a1_difference <- left_join(a1_mean_latency, b1_mean_latency, by = join_by(`private.id`)) %>% mutate(b1_a1_difference = b1 - a1) %>% subset(select = c(`private.id`, b1_a1_difference))

# Calculate difference between mean latency of B2 and A2
a2_mean_latency <- mean_latency_per_block %>% filter(block %in% "A2") %>% rename(a2 = mean_reaction_time)
b2_mean_latency <- mean_latency_per_block %>% filter(block %in% "B2") %>% rename(b2 = mean_reaction_time)
b2_a2_difference <- left_join(a2_mean_latency, b2_mean_latency, by = join_by(`private.id`)) %>% mutate(b2_a2_difference = b2 - a2) %>% subset(select = c(`private.id`, b2_a2_difference))

# Join in one table on participant ID
mean_latency_differences <- left_join(b1_a1_difference, b2_a2_difference)
mean_latency_differences
```

# Step 8

Do the big function

```{r}
participant_stats_per_block <- step6 %>% group_by(`private.id`, block) %>% summarise(mean = mean(`reaction.time`), sd = sd(`reaction.time`), count = n())
a1_stats <- participant_stats_per_block %>% filter(block %in% "A1") %>% rename(a1_mean = mean, a1_sd = sd, a1_count = count) %>% subset(select = -c(block))
a2_stats <- participant_stats_per_block %>% filter(block %in% "A2") %>% rename(a2_mean = mean, a2_sd = sd, a2_count = count) %>% subset(select = -c(block))
b1_stats <- participant_stats_per_block %>% filter(block %in% "B1") %>% rename(b1_mean = mean, b1_sd = sd, b1_count = count) %>% subset(select = -c(block))
b2_stats <- participant_stats_per_block %>% filter(block %in% "B2") %>% rename(b2_mean = mean, b2_sd = sd, b2_count = count) %>% subset(select = -c(block))
all_stats <- left_join(left_join(left_join(a1_stats, a2_stats), b1_stats), b2_stats)

inclusive_sd <- mutate(
  all_stats,
  .keep = "none",
  sd1 = sqrt((((a1_count - 1) * a1_sd^2 + (b1_count - 1) * b1_sd^2) + ((a1_count + b1_count) * ((a1_mean - b1_mean)^2) / 4)) / (a1_count + b1_count - 1)),
  sd2 = sqrt((((a2_count - 1) * a2_sd^2 + (b2_count - 1) * b2_sd^2) + ((a2_count + b2_count) * ((a2_mean - b2_mean)^2) / 4)) / (a2_count + b2_count - 1))
)
```

# Step 9

Compute (b1-a1) / sd1 and (b2-a2) / sd2

```{r}
step9 <- left_join(mean_latency_differences, inclusive_sd) %>% mutate(.keep = "none", x = b1_a1_difference / sd1, y = b2_a2_difference / sd2)
step9
```

# Step 10

Take the average of the two quotients computed in step 9

```{r}
step10 <- step9 %>% mutate(.keep = "none", result = (x + y) / 2) %>% arrange(desc(result))
step10
boxplot(data = step10, x=step10$result, xlab="D Score", ylim=c(-2, 2))
```

# Export csv

```{r}
write.csv(step10, "data/iat_scores.csv")
```

