---
title: "trial_markdown"
output:
  html_document: default
  word_document: default
date: "2025-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matching


```{r unmatched}
install.packages("MatchIt@4.7.2")
#install.packages("optmatch@0.10.8")

library("MatchIt")

pz_data <- read.csv("data/data_unmatched.csv")

pz_data$education <- factor(pz_data$education, levels=c('None specified', 'Secondary school', 'Further education', 'Undergraduate degree', 'Postgraduate degree'))
pz_data$donation_habits <- factor(pz_data$donation_habits, levels=c('Never', 'Yearly', 'Once every 6 months', 'Once every 3 months', 'Monthly'))

unmatched <- matchit(player ~ age + gender + education + nr + donation_habits,
                     data = pz_data,
                     method = NULL,
                     distance = "glm")

summary(unmatched)
```



## Nearest Neighbour

Method: nearest neighbour
Distance = glm
Order = closest
Caliper = none, 1, 0.75, or 0.5

```{r method1, echo = FALSE}
matched1 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "closest", 
                    caliper = NULL)

summary(matched1, un = TRUE)

plot(summary(matched1))

plot(matched1, type = "jitter", interactive = FALSE)

plot(matched1, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method2, echo = FALSE}
matched2 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "closest", 
                    caliper = 1)

summary(matched2, un = TRUE)

plot(summary(matched2))

plot(matched2, type = "jitter", interactive = FALSE)

plot(matched2, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method3, echo = FALSE}
matched3 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "closest", 
                    caliper = 0.75)

summary(matched3, un = TRUE)

plot(summary(matched3))

plot(matched3, type = "jitter", interactive = FALSE)

plot(matched3, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method4, echo = FALSE}
matched4 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "closest", 
                    caliper = 0.5)

summary(matched4, un = TRUE)

plot(summary(matched4))

plot(matched4, type = "jitter", interactive = FALSE)

plot(matched4, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



## Nearest Neighbour

Method: nearest neighbour
Distance = glm
Order = farthest
Caliper = none, 1, 0.75, or 0.5

```{r method5, echo = FALSE}
matched5 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "farthest", 
                    caliper = NULL)

summary(matched5, un = TRUE)

plot(summary(matched5))

plot(matched5, type = "jitter", interactive = FALSE)

plot(matched5, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method6, echo = FALSE}
matched6 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "farthest", 
                    caliper = 1)

summary(matched6, un = TRUE)

plot(summary(matched6))

plot(matched6, type = "jitter", interactive = FALSE)

plot(matched6, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method7, echo = FALSE}
matched7 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "farthest", 
                    caliper = 0.75)

summary(matched7, un = TRUE)

plot(summary(matched7))

plot(matched7, type = "jitter", interactive = FALSE)

plot(matched7, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method8, echo = FALSE}
matched8 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = "farthest", 
                    caliper = 0.5)

summary(matched8, un = TRUE)

plot(summary(matched8))

plot(matched8, type = "jitter", interactive = FALSE)

plot(matched8, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Nearest Neighbour

Method: nearest neighbour
Distance = glm
Order = none
Caliper = none, 1, 0.75, or 0.5

```{r method9, echo = FALSE}
matched9 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = NULL, 
                    caliper = NULL)

summary(matched9, un = TRUE)

plot(summary(matched9))

plot(matched9, type = "jitter", interactive = FALSE)

plot(matched9, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method10, echo = FALSE}
matched10 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = NULL, 
                    caliper = 1)

summary(matched10, un = TRUE)

plot(summary(matched10))

plot(matched10, type = "jitter", interactive = FALSE)

plot(matched10, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method11, echo = FALSE}
matched11 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = NULL, 
                    caliper = 0.75)

summary(matched11, un = TRUE)

plot(summary(matched11))

plot(matched11, type = "jitter", interactive = FALSE)

plot(matched11, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method12, echo = FALSE}
matched12 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits,
                    m.order = NULL, 
                    caliper = 0.5)

summary(matched12, un = TRUE)

plot(summary(matched12))

plot(matched12, type = "jitter", interactive = FALSE)

plot(matched12, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



## Nearest Neighbour

Method: nearest neighbour
Distance = mahalanobis
Order = closest
Caliper = none, 1, 0.75, or 0.5


```{r method13, echo = FALSE}
matched13 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "closest", 
                    caliper = NULL)

summary(matched13, un = TRUE)

plot(summary(matched13))

## plot(matched13, type = "jitter", interactive = FALSE)

plot(matched13, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method14, echo = FALSE}
matched14 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "closest", 
                    caliper = c(age = 1, nr = 1))

summary(matched14, un = TRUE)

plot(summary(matched14))

## plot(matched14, type = "jitter", interactive = FALSE)

plot(matched14, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method15, echo = FALSE}
matched15 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "closest", 
                    caliper = c(age = .75, nr = 0.75))

summary(matched15, un = TRUE)

plot(summary(matched15))

## plot(matched15, type = "jitter", interactive = FALSE)

plot(matched15, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method16, echo = FALSE}
matched16 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "closest", 
                    caliper = c(age = .5, nr = 0.5))

summary(matched16, un = TRUE)

plot(summary(matched16))

## plot(matched16, type = "jitter", interactive = FALSE)

plot(matched16, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Nearest Neighbour

Method: nearest neighbour
Distance = mahalanobis
Order = farthest
Caliper = none, 1, 0.75, or 0.5

```{r method17, echo = FALSE}
matched17 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "farthest", 
                    caliper = NULL)

summary(matched17, un = TRUE)

plot(summary(matched17))

## plot(matched17, type = "jitter", interactive = FALSE)

plot(matched17, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method18, echo = FALSE}
matched18 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "farthest", 
                    caliper = c(age = 1, nr = 1))

summary(matched18, un = TRUE)

plot(summary(matched18))

## plot(matched18, type = "jitter", interactive = FALSE)

plot(matched18, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method19, echo = FALSE}
matched19 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "farthest", 
                    caliper = c(age = .75, nr = 0.75))

summary(matched19, un = TRUE)

plot(summary(matched19))

## plot(matched19, type = "jitter", interactive = FALSE)

plot(matched19, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method20, echo = FALSE}
matched20 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = "farthest", 
                    caliper = c(age = 0.5, nr = 0.5))

summary(matched20, un = TRUE)

plot(summary(matched20))

## plot(matched20, type = "jitter", interactive = FALSE)

plot(matched20, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Nearest Neighbour

Method: nearest neighbour
Distance = mahalanobis
Order = none
Caliper = none, 1, 0.75, or 0.5

```{r method21, echo = FALSE}
matched21 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = NULL, 
                    caliper = NULL)

summary(matched21, un = TRUE)

plot(summary(matched21))

## plot(matched21, type = "jitter", interactive = FALSE)

plot(matched21, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method22, echo = FALSE}
matched22 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = NULL, 
                    caliper = c(age = 1, nr = 1))

summary(matched22, un = TRUE)

plot(summary(matched22))

## plot(matched22, type = "jitter", interactive = FALSE)

plot(matched22, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method23, echo = FALSE}
matched23 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = NULL, 
                    caliper = c(age = 0.75, nr = 0.75))

summary(matched23, un = TRUE)

plot(summary(matched23))

## plot(matched23, type = "jitter", interactive = FALSE)

plot(matched23, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```



```{r method24, echo = FALSE}
matched24 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "nearest", 
                    distance = "mahalanobis", 
                    m.order = NULL, 
                    caliper = c(age = .5, nr = 0.5))

summary(matched24, un = TRUE)

plot(summary(matched24))

## plot(matched24, type = "jitter", interactive = FALSE)

plot(matched24, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Optimal Pair

Method: optimal pair
Distance = glm
Order = none
Caliper = none

```{r method25, echo = FALSE}
matched25 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "optimal", 
                    distance = "glm", 
                    mahvars = ~ age + gender + education + nr + donation_habits)

summary(matched25, un = TRUE)

plot(summary(matched25))

plot(matched25, type = "jitter", interactive = FALSE)

plot(matched25, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Optimal Pair

Method: optimal pair
Distance = mahalanobis
Order = none
Caliper = none

```{r method26, echo = FALSE}
matched26 <- matchit(player ~ age + gender + education + nr + donation_habits,
                    data = pz_data, method = "optimal", 
                    distance = "mahalanobis")

summary(matched26, un = TRUE)

plot(summary(matched26))

## plot(matched26, type = "jitter", interactive = FALSE)

plot(matched26, type = "density", interactive = FALSE,
     which.xs = ~age + gender + education + nr + donation_habits)
```


## Specification number 6 seems best.

```{r exported matched data, eco = FALSE}
get_matches(matched6)
final_data <- get_matches(matched6)
write.csv(final_data, "data/data_matched.csv")
```
